Reproducible Research
=====================
Project 1
MZ

**Total number of steps each day**
Transform data 
```{r}
setwd("C:\\users\\zhuangmg\\Coursera\\Reproducible Research\\project 1")
ActivityTable01<-read.csv("activity.csv")
ActivityTable02<-aggregate(ActivityTable01$steps,by=list(ActivityTable01$date),FUN=sum,na.rm=TRUE)
names(ActivityTable02)<-c("date","TotalSteps")
print(ActivityTable02)
```

**Total number of steps each day**
Construct histogram of total number of steps each day
```{r}
hist(ActivityTable02$TotalSteps,main="Histogramof Total Steps of Each Day",xlab="Total Steps")
```

**Total number of steps each day**
Calculate mean & median
```{r}
mean(ActivityTable02$TotalSteps)
median(ActivityTable02$TotalSteps)
```

**Average daily activity pattern**
Transform data
```{r}
ActivityTable03<-aggregate(ActivityTable01$steps,by=list(ActivityTable01$interval),FUN=mean,na.rm=TRUE)
names(ActivityTable03)<-c("interval","AverageSteps")
print(ActivityTable03)
```

**Average daily activity pattern**
Plot time series data 
```{r}
plot(ActivityTable03$interval,ActivityTable03$AverageSteps,type = "l",main="Daily Activity Pattern",xlab="Time Interval (5 min)", ylab="Average Steps")
```

**Average daily activity pattern**
Find the interval that contains the maximum number of steps
```{r}
ActivityTable03[which.max(ActivityTable03$AverageSteps),]
```

**Imputing missing values**
calculate the total number of rows with missing values
```{r}
nrow(ActivityTable01[!complete.cases(ActivityTable01),])
```

**Imputing missing values**
Impute missing values using the mean value of that interval and create a new dataset
```{r}
fillNA <- numeric()
for (i in 1:nrow(ActivityTable01)) {
    obs <- ActivityTable01[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(ActivityTable03, interval == obs$interval)$AverageSteps
    } else {
        steps <- obs$steps
    }
    fillNA <- c(fillNA, steps)
}
ActivityTable04=ActivityTable01
ActivityTable04$steps=fillNA
nrow(ActivityTable04[!complete.cases(ActivityTable04),])
```
Now ther is no missing values in the dataset

**Imputing missing values**
Transform data, construct histogram of total number of steps each day, and compare with previous histogram
```{r}
ActivityTable05<-aggregate(ActivityTable04$steps,by=list(ActivityTable04$date),FUN=sum,na.rm=TRUE)
names(ActivityTable05)<-c("date","TotalSteps")
hist(ActivityTable05$TotalSteps,main="Histogramof Total Steps of Each Day",xlab="Total Steps")
```

This second histogram is different because of the imputation of missing values.

**Imputing missing values**
Calculate mean & median and compare with previous values
```{r}
mean(ActivityTable05$TotalSteps)
median(ActivityTable05$TotalSteps)
```
The mean and median are also different from previous values because of the imputation of missing values.

**Differences between weekdays and weekends**
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" 
```{r}
ActivityTable04$date <- as.Date(ActivityTable04$date, "%Y-%m-%d")
ActivityTable04$weekdays<-weekdays(ActivityTable04$date)
ActivityTable04weekdays<-as.factor(ActivityTable04$weekdays)
ActivityTable04$daylevel=ActivityTable04$weekdays
for (i in 1:nrow(ActivityTable04)){
      if (ActivityTable04$weekdays[i]=="Saturday"){
      ActivityTable04$daylevel[i]<-"weekend"
    } else if (ActivityTable04$weekdays[i]=="Sunday"){
      ActivityTable04$daylevel[i]<-"weekend"
    } else {
      ActivityTable04$daylevel[i]<-"weekday"
    }
}
ActivityTable06 <- aggregate(steps~interval+daylevel,FUN=mean,data=ActivityTable04)
names(ActivityTable06)<-c("interval","daylevel","AverageSteps")
```

**Differences between weekdays and weekends**
```{r}
library(lattice)
xyplot(AverageSteps ~ interval | daylevel,ActivityTable06, type = "l", layout = c(1, 2), main="Daily Activity Pattern:weekend vs. weekday",xlab="Time Interval (5 min)", ylab="Average Steps")
```

The patterns of weekday and weekend are slightly different.